### 2021-09-23 Research Paper/figures/tex_files/gan_visualization.tex

%\definecolor{offwhite}{RGB}{255, 255, 243}  
%\pagecolor{offwhite}
\begin{tikzpicture}[box/.style={draw,thick,minimum width=5em,minimum
    height=10em},arj/.style={semithick,-latex}]
  \begin{scope}[start chain=A going right,oc/.style={join=by arj,on chain},
    font=\sffamily,node distance=1.75em]

   \node(color data)[outer sep=0pt,inner sep=0pt,label=Ground Truth] at (1.2,1) {\includegraphics[width=2cm]{figures/tex_files/ground_truth.jpg}};
   \node(discriminator)[box,minimum height=7em, minimum width=7em,anchor=west] at (4,-2) {\scriptsize{Discriminator}}; 
   \draw[->] (color data) -| (discriminator);

   \node(generator)[box, minimum height=7em, minimum width=7em] at (-2,-2){\scriptsize{Generator}};
   \node(input image)[outer sep=0pt,inner sep=0pt,label=Grayscale Image] at (-5.5,-2) {\includegraphics[width=2cm]{figures/tex_files/input_grayscale.jpg}};
   \draw[->] (input image) -- (generator);

   \node(generator func)[outer sep=0pt,inner sep=0pt,label=Generated Image] at (1.2,-2){\includegraphics[width=2cm]{figures/tex_files/generated_output.jpg}};
   \draw[<-] (generator func) -- (generator);
   \draw[->] (generator func) -- (discriminator);

   \node(discriminator func)[box, minimum height = 2em, minimum width = 1em] at (8, -2){\textit{\scriptsize{p = 0.82}}};

   \draw[->] (discriminator) -- (discriminator func);

   \node(cost)[box, minimum height = 2em, minimum width = 3em] at (10, -2){\scriptsize{cost}};

   \draw[->] (discriminator func) -- (cost);

   \draw[dashed, ->] (cost) -- (12, -2) -- node[above left=4em]{\scriptsize{loss of discriminator}} (12, 0) -| ($(discriminator) +(0.5,1.3)$);

   \draw[dashed, ->] (12,-2) -- node[below left =4em]{\scriptsize{loss of Generator}} (12, -4) -| (generator);

  \end{scope}
%   \begin{scope}[shift = {{($(input.south)+(0cm, -2cm)$)}}, start chain=A going right, oc/.style={join=by arj, on chain},
%     font=\sffamily, node distance=1.75em]
%     \node[box,oc]{something};
%   \end{scope}
\end{tikzpicture}